Quality Definitions of Qualities

### 2019-11-06 StateOfPractice/QDefOfQualities/QDefOfQualities.tex
Reformat tex file for 80 character width

\subsection{Installability}
\begin{itemize}
\item{Definition 1:} A measure of the ease of installation.\\
\item{Definition 2:} Installability is related to the amount of human effort required to install a software in a designated environment \cite{berander2005software}. \\
\item{Definition 3:} ``The capability of a software product to be installed in a specified environment'' \cite{berander2005software}.\\
\item{Definition 4:}  ``The efficacy of the installation, uninstallation or reinstallation process of a software product in a specified enviroment'' (\cite{iso201017043})\\
\item {\textbf{Proposed definition}:} (Definition 2)
\end{itemize}

@@

\subsection{Verifiability}
\begin{itemize}
\item{Definition 1:} A software is verifiable if its properties can be verified
  (\cite{GhezziEtAl2003}).
\item{Definition 2:} Attributes of software that relate to the effort needed for
  validating the modified software (\cite{berander2005software}).
\item{Definition 3:}) The degree to which a system or component facilitates the
  establishment of test criteria and the performance of tests to determine
  whether those criteria have been met (\cite{IEEEStdGlossarySET1990}).
\item{Definition 4:} The degree to which a requirement is stated in terms that
  permit establishment of test criteria and performance
  (\cite{IEEEStdGlossarySET1990}).
\item{Definition 5:} Ease of performing testing on a software product or system
  (\cite{IEEEStdGlossarySET1990}).
\item{\textbf{Proposed definition}:} Combo of definitions 4 and 5 ?
\end{itemize}
\subsubsection{Discussion}
Verifiability involves ``solving the equations right''~\cite[p.~23]{Roache1998};
it benefits from rational documentation that systematically shows, with explicit
traceability, how the governing equations are transformed into code.
\oo{Verifiability is sometimes referred to as testability, so I culled some
  testability definitions here}

@@

\subsection{Reliability}
\begin{itemize}
\item{Definition 1:} The probability that the software will operate as expected over a specified time interval (\cite{GhezziEtAl2003}).\\
\item{Definition 2:} A set of attributes that relate to the capability of software to maintain its level of performance under stated conditions for a stated period of time (\cite{berander2005software}).\\
\item{Definition 3:} “The capability of the software product to maintain a specified level of performance when used under specified conditions (\cite{international2001iso}).\\
\item{Definition 4:} Code possesses the characteristic reliability to the extent that it can be expected to perform its intended functions satisfactorily (\cite{boehm2007software})\\
\item{Definition 5:} A concern encompassing correctness and robustness (\cite{meyer1988object}).\\
\item{Proposed definition}\\
\end{itemize}

\subsubsection{Discussion}
Reliability is a critical quality for scientific software, since the results of
computations are meaningless, if they are not dependable.  Reliability is
closely tied to verifiability, since the key quality to verify is reliability,
while the act of verification itself improves reliability.\\
Reliability models can be used to predict reliability of a software product. For
example measuring Mean Time to Fail (MTTF) can be a good measure of reliability
\cite{berander2005software}

@@

Nielsen and (separately) Schneidermann have defined usability as part of usefulness and
is composed of
\begin{itemize}
\item Learnability: How easy is it for users to accomplish basic tasks the
  first time they encounter the design?
\item Efficiency: Once users have learned the design, how quickly can they perform tasks?
\item Memorability: When users return to the design after a period of not using
  it, how easily can they re-establish proficiency?
\item Errors: How many errors do users make, how severe are these errors, and
  how easily can they recover from the errors?
\item Satisfaction: How pleasant is it to use the design?
\end{itemize}
In that context, it makes sense to separate \emph{usefulness} into
\emph{usability} (purely an interface concern) and \emph{utility} (in the economics
sense of the word).

@@

The Interaction Design Foundation~\url{https://www.interaction-design.org/literature/topics/usability}
further lists the following desirable outcomes:
\begin{enumerate}
\item It should be easy for the user to become familiar with and competent in using
the user interface during the first contact with the website. For example, if a
travel agent’s website is a well-designed one, the user should be able to move
through the sequence of actions to book a ticket quickly.
\item It should be easy for users to achieve their objective through using the
website. If a user has the goal of booking a flight, a good design will guide
him/her through the easiest process to purchase that ticket.
\item It should be easy to recall the user interface and how to use it on
subsequent visits. So, a good design on the travel agent’s site means the user
should learn from the first time and book a second ticket just as easily.
\end{enumerate}

@@

\subsection{Maintainability}
\begin{defn} \label{MaintainabilityDefnSelected1}
The ease with which a software system or component can be modified to correct faults, improve performance or other attributes, or adapt to a changed environment \citep{IEEEStdGlossarySET1990}. 
\end{defn}\
\begin{defn}
ISO/IEC 25010 refers to maintainability as the degree of effectiveness and efficiency with which a product or system can be modified by the intended maintainers \citep{ISO/IEC25010}.
\end{defn}
\begin{defn}
Effort required to locate and fix an error in a program
\citep{pressman2005software}.
\end{defn}
\begin{defn}
A set of attributes that bear on the effort needed to make specified modifications (which may include corrections, improvements, or adaptations of software to environmental changes and changes in the requirements and functional specifications)\citep{pfleeger2006software}.
\end{defn}
\begin{defn}
We will view maintainability as two separate qualities: repairability and evolvability. Software is repairable if it allows the fixing of defects; it is evolvable if it allows changes that enable it to satisfy new requirements \citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn} \label{MaintainabilityDefnSelected2}
Code possesses the characteristic of maintainability to the extent that it facilitates updating to satisfy new requirements or to correct deficiencies \citep{boehm2007software}.
\end{defn}

@@

\subsection{Interoperability}
\noindent \textbf{Chosen definition:} Def \ref{InteroperabilitySelected}
rephrased: Interoperability is the ability of different information systems,
devices and applications (‘systems’) to access, exchange, integrate and
cooperatively use data in a coordinated manner, within and across
organizational, regional and national boundaries, to provide timely and seamless
portability of information.

@@

\subsection{Productivity}
\noindent \textbf{Chosen definition:} The first sentence of Def
\ref{ProductivitySelected}: Productivity is the amount of output (what is
produced) per unit of input used.
\begin{defn}
The best definition of the productivity of a process is
\[\text{Productivity} = \dfrac{\text{Outputs produced by the
process}}{\text{Inputs consumed by the process}}\]
Defining inputs. For the software process, providing a meaningful definition of
inputs is a nontrivial but generally workable problem. Inputs to the software
process generally comprise labor, computers, supplies, and other support
facilities and equipment. Defining outputs. The big problem in defining software
productivity is defining outputs. Here we find a defining delivered source
instructions (DSI) or lines of code as the output of the software process is
totally inadequate, and they argue that there are a number of deficiencies in
using DSI. However, most organizations doing practical productivity measurement
still use DSI as their primary metric \citep{Boehm1987}.
\end{defn}
\begin{defn}
\label{ProductivitySelected}
Productivity is the amount of output (what is produced) per unit of input
used.If we can measure the size of the software product and the effort required
to develop the product, we have:
\begin{align}
\text{productivity} = \text{size}/\text{effort}
\end{align}
Equation (1) assumes that size is the output of the software production process
and effort is the input to the process. This can be contrasted with the
viewpoint of software cost models where we use size as an independent variable
(i.e., an input) to predict effort which is treated as an output. Equation (1)
is simple to operationalize if we have a single dominant size measure, for
example, product size measured in lines of code \citep{Kitchenham2004}.
\end{defn}

@@

\subsection{Completeness}
\noindent \textbf{Chosen definition:} The first sentence of Def
\ref{CompletenessSelected}: A specification is complete to the extent that all
of its parts are present and each part is fully developed.

@@

\subsection{Consistency}
\noindent \textbf{Chosen definition:} Def \ref{ConsistencySelected}.

>>>>

\subsection{Installability}
\begin{itemize}
\item{Definition 1:} A measure of the ease of installation.
\item{Definition 2:} Installability is related to the amount of human effort
  required to install a software in a designated environment
  \cite{berander2005software}.
\item{Definition 3:} ``The capability of a software product to be installed in a
  specified environment'' \cite{berander2005software}.
\item{Definition 4:} ``The efficacy of the installation, uninstallation or
  reinstallation process of a software product in a specified enviroment''
  (\cite{iso201017043})
\item {\textbf{Proposed definition}:} (Definition 2)
\end{itemize}

@@

\begin{itemize}
\item{Definition 1:} A software is verifiable if its properties can be verified
  (\cite{GhezziEtAl2003}).
\item{Definition 2:} Attributes of software that relate to the effort needed for
  validating the modified software (\cite{berander2005software}).
\item{Definition 3:}) The degree to which a system or component facilitates the
  establishment of test criteria and the performance of tests to determine
  whether those criteria have been met (\cite{IEEEStdGlossarySET1990}).
\item{Definition 4:} The degree to which a requirement is stated in terms that
  permit establishment of test criteria and performance
  (\cite{IEEEStdGlossarySET1990}).
\item{Definition 5:} Ease of performing testing on a software product or system
  (\cite{IEEEStdGlossarySET1990}).
\item{\textbf{Proposed definition}:} Combo of definitions 4 and 5 ?
\end{itemize}

Verifiability involves ``solving the equations right''~\cite[p.~23]{Roache1998};
it benefits from rational documentation that systematically shows, with explicit
traceability, how the governing equations are transformed into code.
\oo{Verifiability is sometimes referred to as testability, so I culled some
  testability definitions here}

@@

\subsection{Reliability}

\begin{itemize}
\item{Definition 1:} The probability that the software will operate as expected
  over a specified time interval (\cite{GhezziEtAl2003}).
\item{Definition 2:} A set of attributes that relate to the capability of
  software to maintain its level of performance under stated conditions for a
  stated period of time (\cite{berander2005software}).
\item{Definition 3:} “The capability of the software product to maintain a
  specified level of performance when used under specified conditions
  (\cite{international2001iso}).
\item{Definition 4:} Code possesses the characteristic reliability to the extent
  that it can be expected to perform its intended functions satisfactorily
  (\cite{boehm2007software})
\item{Definition 5:} A concern encompassing correctness and robustness
  (\cite{meyer1988object}).
\item{Proposed definition}
\end{itemize}

Reliability is a critical quality for scientific software, since the results of
computations are meaningless, if they are not dependable.  Reliability is
closely tied to verifiability, since the key quality to verify is reliability,
while the act of verification itself improves reliability.

Reliability models can be used to predict reliability of a software product. For
example measuring Mean Time to Fail (MTTF) can be a good measure of reliability
\cite{berander2005software}.

@@

Nielsen and (separately) Schneidermann have defined usability as part of usefulness and
is composed of:
\begin{itemize}
\item Learnability: How easy is it for users to accomplish basic tasks the
  first time they encounter the design?
\item Efficiency: Once users have learned the design, how quickly can they perform tasks?
\item Memorability: When users return to the design after a period of not using
  it, how easily can they re-establish proficiency?
\item Errors: How many errors do users make, how severe are these errors, and
  how easily can they recover from the errors?
\item Satisfaction: How pleasant is it to use the design?
\end{itemize}
In that context, it makes sense to separate \emph{usefulness} into
\emph{usability} (purely an interface concern) and \emph{utility} (in the economics
sense of the word).

@@

The Interaction Design Foundation~\url{https://www.interaction-design.org/literature/topics/usability}
further lists the following desirable outcomes:

\begin{enumerate}
\item It should be easy for the user to become familiar with and competent in using
the user interface during the first contact with the website. For example, if a
travel agent’s website is a well-designed one, the user should be able to move
through the sequence of actions to book a ticket quickly.
\item It should be easy for users to achieve their objective through using the
website. If a user has the goal of booking a flight, a good design will guide
him/her through the easiest process to purchase that ticket.
\item It should be easy to recall the user interface and how to use it on
subsequent visits. So, a good design on the travel agent’s site means the user
should learn from the first time and book a second ticket just as easily.
\end{enumerate}

@@

\subsection{Maintainability}

\begin{defn} \label{MaintainabilityDefnSelected1}
The ease with which a software system or component can be modified to correct
faults, improve performance or other attributes, or adapt to a changed
environment \citep{IEEEStdGlossarySET1990}. 
\end{defn}\
\begin{defn}
ISO/IEC 25010 refers to maintainability as the degree of effectiveness and
efficiency with which a product or system can be modified by the intended
maintainers \citep{ISO/IEC25010}.
\end{defn}
\begin{defn}
Effort required to locate and fix an error in a program
\citep{pressman2005software}.
\end{defn}
\begin{defn}
A set of attributes that bear on the effort needed to make specified
modifications (which may include corrections, improvements, or adaptations of
software to environmental changes and changes in the requirements and functional
specifications)\citep{pfleeger2006software}.
\end{defn}
\begin{defn}
We will view maintainability as two separate qualities: repairability and
evolvability. Software is repairable if it allows the fixing of defects; it is
evolvable if it allows changes that enable it to satisfy new requirements
\citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn} \label{MaintainabilityDefnSelected2}
Code possesses the characteristic of maintainability to the extent that it
facilitates updating to satisfy new requirements or to correct deficiencies
\citep{boehm2007software}.
\end{defn}

@@

\subsection{Interoperability}

\noindent \textbf{Chosen definition:} Def \ref{InteroperabilitySelected}
rephrased: Interoperability is the ability of different information systems,
devices and applications (‘systems’) to access, exchange, integrate and
cooperatively use data in a coordinated manner, within and across
organizational, regional and national boundaries, to provide timely and seamless
portability of information.

@@

\subsection{Productivity}

\noindent \textbf{Chosen definition:} The first sentence of Def
\ref{ProductivitySelected}: Productivity is the amount of output (what is
produced) per unit of input used.
\begin{defn}
The best definition of the productivity of a process is
\[\text{Productivity} = \dfrac{\text{Outputs produced by the
process}}{\text{Inputs consumed by the process}}\]
Defining inputs. For the software process, providing a meaningful definition of
inputs is a nontrivial but generally workable problem. Inputs to the software
process generally comprise labor, computers, supplies, and other support
facilities and equipment. Defining outputs. The big problem in defining software
productivity is defining outputs. Here we find a defining delivered source
instructions (DSI) or lines of code as the output of the software process is
totally inadequate, and they argue that there are a number of deficiencies in
using DSI. However, most organizations doing practical productivity measurement
still use DSI as their primary metric \citep{Boehm1987}.
\end{defn}
\begin{defn}
\label{ProductivitySelected}
Productivity is the amount of output (what is produced) per unit of input
used.If we can measure the size of the software product and the effort required
to develop the product, we have:
\begin{align}
\text{productivity} = \text{size}/\text{effort}
\end{align}
Equation (1) assumes that size is the output of the software production process
and effort is the input to the process. This can be contrasted with the
viewpoint of software cost models where we use size as an independent variable
(i.e., an input) to predict effort which is treated as an output. Equation (1)
is simple to operationalize if we have a single dominant size measure, for
example, product size measured in lines of code \citep{Kitchenham2004}.
\end{defn}

@@

\subsection{Completeness}

\noindent \textbf{Chosen definition:} The first sentence of Def
\ref{CompletenessSelected}: A specification is complete to the extent that all
of its parts are present and each part is fully developed.

@@

\subsection{Consistency}

\noindent \textbf{Chosen definition:} Def \ref{ConsistencySelected}.
